---
title: "Additional excercises - data wrangling"
output: html_notebook
---

Following excercises and data description below are taken from [R for Data Science by Hadley Wickham](https://r4ds.had.co.nz). Deffo check it out!

The dataset comes from `nycflights13` package, the `flights` data frame contains all 336,776 flights that departed from New York City in 2013. The data comes from the US Bureau of Transportation Statistics, and is documented in `?flights`

```{r}
if (!require("nycflights13", character.only = TRUE, 
             warn.conflicts = FALSE, quietly = TRUE)) {
  print("Installing nycflights13 package.")
  install.packages("nycflights13")
}
  
library(nycflights13)
flights
```

Find all flights that  
  * Had an arrival delay of two or more hours  
  * Flew to Houston (IAH or HOU)  
  * Were operated by United, American, or Delta  
  * Departed in summer (July, August, and September)  
  * Arrived more than two hours late, but didn’t leave late  
  * Were delayed by at least an hour, but made up over 30 minutes in flight  
  * Departed between midnight and 6am (inclusive)  

```{r}
# your code goes here
```

Another useful dplyr filtering helper is `between()`. What does it do? Can you use it to simplify the code needed to answer the previous challenges?

```{r}
# your code goes here
```

How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
# your code goes here
```


Which flights travelled the longest? Which travelled the shortest?

```{r}
# your code goes here
```

What does the one_of() function do? Why might it be helpful in conjunction with this vector?

```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
# your code goes here
```


Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for `min_rank()`.

```{r}
# your code goes here
```


Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about `flights %>% group_by(carrier, dest) %>% summarise(n())`)

```{r}
# your code goes here
```
