---
title: "RNA-seq analysis"
output: html_notebook
---

```{r setup}
library("DESeq2")
library("ComplexHeatmap")
library("tidyverse")
library("ggsci")
theme_set(theme_classic())
```

# About the data 

[Jividen K, Kedzierska KZ, Yang C, Szlachta K, Ratan A & Paschal BM; _Genomic analysis of DNA repair genes and androgen signaling in prostate cancer_; BMC Cancer 18, 960 (2018)](https://bmccancer.biomedcentral.com/articles/10.1186/s12885-018-4848-x)

The data comes from the paper describing androgen-regulated gene expression. RNA-seq xperiment was performed on 3 prostate cancer cell lines. Two that express androgen receptor (AR): LNCaP and VCaP; and PC3 to which AR expression was introduced by a stable lenti-viral infection of full length AR.  

Relevand *Methods* section:

_For RNA-sequencing experiments, the Qiagen RNeasy kit was used to extract RNA. Library preparation and sequencing was performed by Hudson Alpha. Briefly, RNA integrity and concentration were assessed by a fluorometric assay, indexed libraries were made using the standard polyA method, quality control was used to determine size and concentration, and samples were sequenced using Illumina HiSeq 2500 at a depth of 250 million × 50-bp paired-end reads. Reads were aligned to the hg38 genome (ENSEMBL GRCh38.89) using STAR (release v. 2.5) [14]. Counts were generated using HTSeq (release v. 0.6) [15]_

# Analysis 

## Read in the data 
  
```{r}
count_data <-
  read_csv("./data/GSE120660_PCamerge_hg38.csv") %>%
  column_to_rownames("X1")

sample_sheet <- 
  read_tsv("./data/GSE120660_sample_sheet.tsv")
```

Let us set up consistent coloring for different condition. 

```{r}
cell_lines_colors <- pal_aaas()(length(unique(sample_sheet$cell_lines)))
names(cell_lines_colors) <- unique(sample_sheet$cell_lines)
timepoints_colors <- pal_futurama()(length(unique(sample_sheet$timepoints)))
names(timepoints_colors) <- unique(sample_sheet$timepoints)
treatment_colors <- pal_rickandmorty()(length(unique(sample_sheet$treatment)))
names(treatment_colors) <- unique(sample_sheet$treatment)
```

# Quality check

Do we need all genes?
```{r}
dim(count_data)
```

Are all genes expressed? Are all contributing equally to understanding differences between groups?

```{r}
apply(count_data, 1, var) %>%
  summary()
```

```{r}
count_data_mat <- 
  count_data[ ] 
```

## Samples

Let's chcech the distances between saples. 

```{r}
norm_count <- 
sample_dist <- 

ha = HeatmapAnnotation(df = (sample_sheet %>% 
                               column_to_rownames("sample_names")),
                       col = list(cell_lines = cell_lines_colors,
                                  treatent = treatment_colors,
                                  timepoints = timepoints_colors))


Heatmap(as.matrix(sample_dist), col = viridis::viridis(100), 
        top_annotation = ha)
```

```{r}
pca <- 
var_exp <- pca$sdev^2
var_exp <- round(var_exp / sum(var_exp) * 100, 1)
data.frame(principal_component = paste0("PC", 1:length(var_exp)),
           variance_explained = var_exp,
           stringsAsFactors = FALSE) %>%
  mutate(accumulated_variance = cumsum(variance_explained),
         principal_component = factor(principal_component, 
                                      levels = principal_component)) %>%
  #filter(accumulated_variance < 100) %>%
  ggplot(., aes(x = principal_component, y = variance_explained)) +
  geom_bar(stat = "identity")
```

```{r}
a_pc <- 1
b_pc <- 2
pca_plt <- 
  pca$x %>%
  as.data.frame() %>%
  rownames_to_column("sample_names") %>%
  left_join(sample_sheet) %>% 
  ggplot(., aes_string(x = paste0("PC", a_pc), 
                       y  = paste0("PC", b_pc), 
                       color = "cell_lines",
                       shape = "timepoints")) +
  geom_point() +
  labs(x = paste0("PC", a_pc, " [", var_exp[a_pc], "%]"),
       y = paste0("PC", b_pc, " [", var_exp[b_pc], "%]"),
       title = "PCA") +
  scale_color_aaas()
pca_plt
```


# Differential analysis

Subset for only VCaP, LNCaP and PC-AR andrgen treated at last timepoint. 

