---
title: "R Notebook"
output: html_notebook
---

# Last meeting reminder
## Numbers 

```{r}
my_first_number <- 1
my_second_number = 2
```

## Vectors

```{r}
dbl_vector <- c(1, 2, 3, 4) # double = numeric
int_vector <- c(2L, 4L, 6L, 8L) # integer (integers require less storage space)
chr_vector <- c("this", "is", "a", "character", "vector")
log_vector <- c(TRUE, FALSE, TRUE, T, F) # logical
```

```{r}
this_and_that <- c("this", "that", "this", "this", "that")
fct_vector <- as.factor(this_and_that)
```

```{r}
chr_vector2 <- c(sample(x = this_and_that, 1000000, replace = TRUE), "other")
chr_vector2
fct_vector2 <- as.factor(chr_vector2)
fct_vector2
```


What happens if we put different types of variables in one vector?

```{r}
comb_vect <- c(chr_vector, int_vector)
# comb_vect ?
```

---

## Lists

```{r}
my_list <- list(first_element = log_vector,
                second = chr_vector,
                int_vector)
my_list
```

### Matrices

Matrices are just tables with rows and columns. 

```{r}
my_matrix <- matrix(1:10, nrow = 2, ncol = 5, 
                    dimnames = list(paste0("row_", 1:2),
                                    paste0("column_", 1:5)))
my_matrix
```

### Data frames

```{r}
my_df <- data.frame(column_1 = int_vector,
                    column_2 = c("yes", "no", "yes", "yes"),
                    row.names = paste0("row_", 1:4),
                    stringsAsFactors = FALSE)
my_df
```

## Operations

#### Subsetting 

```{r}
chr_vector[3]
```

```{r}
dbl_vector[1:3]
```

```{r}
chr_vector[c(1,4)]

chr_vector[c(TRUE, FALSE, FALSE, TRUE, FALSE)]
```

```{r}
dbl_vector == c(1, 2, 3)
```

```{r}
chr_vector
```

```{r}
chr_vector[c(TRUE, FALSE)]
```

## Excecise

Given a vector `nums` with numbers from 1 to 20, access the numbers divisible by 5.

```{r}
nums <- 1:20
```

```{r}
# get the numbers divisible by 5
```

We can also create named vectors.

```{r}
my_first_named_vector <- 1:10
names(my_first_named_vector) <- paste0("position_", 1:10)
```

and then access them by names.

```{r}
my_first_named_vector['position_1']
```

```{r}
my_first_named_vector[1]
```


### Lists

Remember the list we created earlier?

```{r}
my_list
```

To access the elements, we can call them by names

```{r}
my_list$second
```

or by partial match :o

```{r}
my_list$f
```

or by positions.

```{r}
my_list[[3]]
```

### Matrices

```{r}
my_matrix
```
By coordinates

```{r}
# get elemnt in row 1 and column 2
my_matrix[1, 2]
```

```{r}
# get column 5
my_matrix[, 5]
# get columns 2 and 4
my_matrix[, c(2,4)]
```

or by names.

```{r}
my_matrix['row_2', 'column_3']
```

When checking conitions, the otput will have the same shape.

```{r}
my_matrix > 1
```

### Data frames

```{r}
my_df
```
By coordinates

```{r}
# get elemnt in row 1 and column 2
my_df[1, 2]
```

```{r}
# get column 2
my_df[, 2]
```


or by names 

```{r}
my_df['row_2', 'column_1']
```

```{r}
my_df$column_2
```


### Excercise

Given the experssion data in a matrix `mat` get the values greater than 80.

```{r}
set.seed(7)
mat <- matrix(runif(50, 0, 100), nrow = 10, ncol = 5, 
                    dimnames = list(paste0("gene_", LETTERS[1:10]),
                                    paste0("sample_", 1:5)))

```

```{r}
# get the values greater than 80
mat[mat > 80]
```

Can you say for which genes and for which samples those values were reached?

# Functions

What is a function?
```{r}
if (!require(tidyverse)) {
  install.packages("tidyverse")
}
library(tidyverse)
df <- data.frame(x = -100:100, 
                 y = (-100:100)^2)

ggplot(df, aes(x, y)) + 
  geom_point() +
  geom_smooth() + 
  theme_classic()
```

```{r}
#' Get the square of a value x
#' @param x vector, values to be squared
#' @return y vector, values squared
x_square <- function(x) {
  x^2
}
```

```{r}
x_square(2)
```

```{r}
x_square(1:5)
```

```{r}
x_ <- -100:100
df <- data.frame(x = x_, 
                 y = x_square(x_))

ggplot(df, aes(x, y)) + 
  geom_point() +
  geom_smooth() + 
  theme_classic()
```

## Excercises

Write a function `my_fun` that will take two arguments - a vector and a nnumber and return the multiplication of this number and this vector.
```{r}

my_fun <- function(vect, numb = 9) {
  numb * vect
}
vect_ <- 1:10
numb <- 7
my_fun(vect = vect_)
```

Write a function `my_fun2` that will take two arguments - a vector and a number and:
* add the number to elemnts at odd postiions of your vector
* mutliply the elements at even positions of your vector
* raise the whole vector to the power of that number

```{r}
my_fun2 <- function(vect, numb) {
  vect[c(TRUE, FALSE)] <- vect[c(TRUE, FALSE)] + numb
  vect[c(FALSE, TRUE)] <- vect[c(FALSE, TRUE)] * numb
  vect ^ numb
}

my_fun2(1:4, 2)
```

## Useful functions?

### Utilty
```{r}
c(1:10, 2, 11)
list(mat_ = mat,
     df_ = my_df)
as.data.frame(mat)
is.numeric(chr_vector)
```

### Maths

```{r}
mean(1:10)
sum(3, 4)
min(dbl_vector)
max(dbl_vector)
```

### Stats

```{r}
TeaTasting <- matrix(c(3, 1, 1, 3),
                     nrow = 2,
                     dimnames = list(Guess = c("Milk", "Tea"),
                                     Truth = c("Milk", "Tea")))
TeaTasting
fisher.test(TeaTasting, alternative = "greater")
```

### Logical

```{r}
log_vector
all(log_vector)
any(log_vector)
```

### Excercise

Remember the expression matrix `mat` from earlier? Calculate `median`, `max` and `min` values for `sample_5` and `gene_C`.

```{r}
# sample_5

```

```{r}
# gene_C

```