---
title: "Introduction to R - Part 2 Functions"
author: "Katarzyna Kedzierska"
date: "14 August 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
---

```{r, include = FALSE}
knitr::opts_chunk$set(message = FALSE,
                      dpi = 300, 
                      out.height = "300px", 
                      out.width = "300px")
```

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
mono_light(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono")
)
```

# Last meeting reminder
## Numbers 

```{r}
my_first_number <- 1
my_second_number = 2
```

## Vectors

```{r}
dbl_vector <- c(1, 2, 3, 4) # double = numeric
int_vector <- c(2L, 4L, 6L, 8L) # integer (integers require less storage space)
chr_vector <- c("this", "is", "a", "character", "vector")
log_vector <- c(TRUE, FALSE, TRUE, T, F) # logical
```

```{r}
this_and_that <- c("this", "that", "this", "this", "that")
fct_vector <- as.factor(this_and_that)
```

---

What happens if we put different types of variables in one vector?

```{r}
comb_vect <- c(chr_vector, int_vector)
# comb_vect ?
```

---

## Lists

```{r}
my_list <- list(first_element = log_vector,
                second = chr_vector,
                int_vector)
my_list
```

---

### Matrices

Matrices are just tables with rows and columns. 

```{r}
my_matrix <- matrix(1:10, nrow = 2, ncol = 5, 
                    dimnames = list(paste0("row_", 1:2),
                                    paste0("column_", 1:5)))
my_matrix
```

---

### Data frames

```{r}
my_df <- data.frame(column_1 = int_vector,
                    column_2 = c("yes", "no", "yes", "yes"),
                    row.names = paste0("row_", 1:4),
                    stringsAsFactors = FALSE)
my_df
```

---

## Operations

#### Subsetting 

```{r}
chr_vector[3]
```

```{r}
dbl_vector[1:3]
```

```{r}
chr_vector[c(1,4)]
```

---
```{r}
dbl_vector == c(1, 2, 3)
```

```{r}
chr_vector
```

```{r}
chr_vector[c(TRUE, FALSE)]
```

---

Excercise!

Given a vector `nums` with numbers from 1 to 20, access the numbers divisible by 5.

```{r}
nums <- 1:20
# nums[]
```

---

Excercise!

Given a vector `nums` with numbers from 1 to 20, access the numbers divisible by 5.

```{r}
nums <- 1:20
nums[nums %% 5 == 0]
nums[c(5, 10, 15, 20)]
```

---
We can also create named vectors.

```{r}
my_first_named_vector <- 1:10
names(my_first_named_vector) <- paste0("position_", 1:10)
```

and then access them by names.

```{r}
my_first_named_vector['position_1']
```

```{r}
my_first_named_vector[1]
```

---

### Lists

Remember the list we created earlier?

```{r}
my_list
```

To access the elements, we can call them by names

```{r}
my_list$second
```

---

or by partial match :o

```{r}
my_list$f
```

or by positions.

```{r}
my_list[[3]]
```

---

### Matrices

```{r}
my_matrix
```
By coordinates

```{r}
# get elemnt in row 1 and column 2
my_matrix[1, 2]
```

```{r}
# get column 5
my_matrix[, 5]
```

---

or by names.

```{r}
my_matrix['row_2', 'column_3']
```

When checking conitions, the otput will have the same shape.

```{r}
my_matrix > 1
```

---
### Data frames

```{r}
my_df
```
By coordinates

```{r}
# get elemnt in row 1 and column 2
my_df[1, 2]
```

```{r}
# get column 2
my_df[, 2]
```

---

or by names 

```{r}
my_df['row_2', 'column_1']
```

```{r}
my_df$column_2
```

---

Excercise.

Given the experssion data in a matrix `mat` get the values greater than 3.

```{r}
set.seed(7)
mat <- matrix(runif(50, 0, 100), nrow = 10, ncol = 5, 
                    dimnames = list(paste0("gene_", LETTERS[1:10]),
                                    paste0("sample_", 1:5)))

mat[mat > 80]
```

Can you say for which genes and for which samples those values were reached?

---

# Functions

What is a function?
```{r}
if (!require(tidyverse)) {
  install.packages("tidyverse")
}
library(tidyverse)
df <- data.frame(x = -100:100, 
                 y = (-100:100)^2)

ggplot(df, aes(x, y)) + 
  geom_point() +
  geom_smooth() + 
  theme_classic()
```

```{r}
#' Get the square of a value x
#' @param x vector, values to be squared
#' @return y vector, values squared
x_square <- function(x) {
  x^2
}
```

---

```{r}
x_square(2)
```

```{r}
x_square(1:5)
```
```{r}
x_ <- -100:100
df <- data.frame(x = x_, 
                 y = x_square(x_))

ggplot(df, aes(x, y)) + 
  geom_point() +
  geom_smooth() + 
  theme_classic()
```
---

## Excercise

Write a function `my_fun` that will take two arguments - a vector and a nnumber and return the multiplication of this number and this vector.
```{r}
my_fun <- function() {
  
}
```

---

## Excercise

Write a function `my_fun` that will take two arguments - a vector and a nnumber and return the multiplication of this number and this vector.
```{r}
my_fun <- function(vec, numb) {
  vec * numb
}
```

---

## Excercise

Write a function `my_fun2` that will take two arguments - a vector and a number and:
* add the number to elemnts at odd postiions of your vector
* mutliply the elements at even positions of your vector
* raise the whole vector to the power of that number

```{r}
my_fun2 <- function() {
  
}
```

---

## Excercise

Write a function `my_fun2` that will take two arguments - a vector and a number and:
* add the number to elemnts at odd postiions of your vector
* mutliply the elements at even positions of your vector
* raise the whole vector to the power of that number

```{r}
my_fun2 <- function(vec, numb) {
  vec2 <- vec
  vec2[vec2 %% 2 == 1] <- vec2[vec2 %% 2 == 1] + numb
  vec2[vec2 %% 2 == 0] <- vec2[vec2 %% 2 == 0] * numb
  vec2 ^ numb
}
```

---
## Useful functions?

### Utilty
```{r}
c(1:10, 2, 11)
list(mat_ = mat,
     df_ = my_df)
as.data.frame(mat)
is.numeric(chr_vector)
```

---
### Maths

```{r}
mean(1:10)
sum(3, 4)
min(dbl_vector)
max(dbl_vector)
```
---
### Stats

```{r}
TeaTasting <- matrix(c(3, 1, 1, 3),
                     nrow = 2,
                     dimnames = list(Guess = c("Milk", "Tea"),
                                     Truth = c("Milk", "Tea")))
TeaTasting
fisher.test(TeaTasting, alternative = "greater")
```

---
### Logical

```{r}
log_vector
all(log_vector)
any(log_vector)
```

---
Excercise

Remember the expression matrix `mat` from earlier? Calculate `median`, `max` and `min` values for `sample_5` and `gene_C`.

```{r}
# sample_5
median(mat[,'sample_5'])
min(mat[,'sample_5'])
max(mat[,'sample_5'])
# gene_C
median(mat['gene_C',])
min(mat['gene_C',])
max(mat['gene_C',])
```

---

There is this nice function that can help us with this.

```{r}
summary(mat)
```

```{r}
summary(t(mat))
```

# Next time

We will learn about the **flow**, i.e.:

* how to easily get the names of genes with expression above median. 
* perform the same operation on each row/column, or element in a vector.
