---
title: "Introduction to R - Part 6"
author: "Katarzyna Kedzierska"
date: "18 September 2019"
output: html_notebook
---

```{r setup}
needed_packages <- c("tidyverse", "ggsci")
for (pkg in needed_packages) {
  if (!require(pkg, quietly = TRUE, character.only = TRUE)) {
    print(paste0("WARNING! Installing ", pkg))
    install.packages(pkg)
  }
  library(pkg, character.only = TRUE)
}
```

# Pipes

And we've touched on dplyr syntax.

Now we know what are pipies and that instead of this:

```{r, eval = FALSE}
leave_house(get_dressed(get_out_of_bed(wake_up(me))))
```

or better yet this:
```{r, eval = FALSE}
woken_up_me <- wake_up(me)
standing_up_me <- get_out_of_bed(woken_up_me)
dressed_me <- get_dressed(standing_up_me)
ready_me <- leave_house(dressed_me)
```

We can write it like this:

```{r, eval = FALSE}
me %>%
  wake_up() %>%
  get_out_of_bed() %>%
  get_dressed() %>%
  leave_house()
```

Source: [Andrew Heiss](https://twitter.com/andrewheiss/status/1173743447171354624)

# Plan for today

* `lapply`, `apply`, ...
* more `tidyverse`

yay!

# Better looping

`lapply`, `apply`, `sapply` and so on, are here to help us write more efficient loops in a more readible way.

But before we start, quick reminder. Given the following matrix:

```{r}
set.seed(7)
mat <- matrix(rnorm(1000), 10, 10)
rownames(mat) <- paste0("gene_", 1:nrow(mat))
colnames(mat) <- paste0("sample_", 1:ncol(mat))
```

get the following:

* `sums_row`, `sum_cols`- sum the rows and columns of the matrix
* `normalized_row` - from each value in the row substract the row' mean and divide the result of substracyion by row' standar deviation

* `sums_row`
```{r}
sums_row <- c()
# your code here
```

* `sums_col`
```{r}
sums_col <- c()
# your code here
```

* `normalized_row`
```{r}
normalized_row <- mat
# your code here
```

# `apply`
Now, let's look how exactly the same thing we can do using `apply`

`apply(array, margin, fun, ...)`

* `array` - array on which we will perform the operation
* `margin` - `1` - perform operation rowwise, `2` - columnwise and `c(1,2)` - both, row and columnwise
* `fun` - function to apply to each row/column

```{r}
sums_row_apply <- apply(mat, 1, sum)
sums_row_apply
```

```{r}
all(sums_row == sums_row_apply)
```

* `sums_col`
```{r, eval = FALSE}
sums_col_apply <- apply()
```

```{r}
all(sums_col == sums_col_apply)
```

We can also define more complex functions:

* `normalized_row`
```{r}
normalized_row_apply <- apply(mat, 1, function(row) (row - mean(row)) / sd(row))
```

```{r}
all(normalized_row_apply == normalized_row)
```

# `lapply`

`lapply(x, fun, ...)`

* `x` - a list over which it will iterate
* `fun` - function to be applied to each item in the list

```{r}
list_in <- list(test = 1:10, train = 10:20, teach = 20:30)
out <- lapply(list_in, median)
out
```

---
# `lapply` vs. `sapply`
```{r}
lout <- lapply(colnames(mat), function(colname) {
 median(mat[, colname])
})

head(lout, n = 3)
```

```{r}
sout <- sapply(colnames(mat), function(colname) {
 median(mat[, colname])
})
sout
```

# Others

* `vapply(x, fun, fun.value)` - perform an operation on each element in the vector
# More tidyverse

```{r}
library(tidyverse)
data(starwars)
starwars
```

## Revision

### summarise

```{r}
starwars %>% 
  summarise(median_height = median(height),
            median_mass = median(mass))
```

### group_by

```{r}
starwars %>%
  group_by(gender) %>%
  summarise(median_height = median(height),
            median_mass = median(mass))
```

### mutate

```{r}
starwars %>%
  mutate(height_in = height / 2.54)
```

### filter

```{r}
starwars %>%
  group_by(gender) %>%
  filter(!is.na(gender)) %>%
  summarise(median_height = median(height),
            median_mass = median(mass))
```

### top_n

```{r}
starwars %>%
  group_by(gender) %>%
  top_n(1, height)
```


```{r}
vapply(1:ncol(mat), function(i) sd(mat[,i]), c(1))
```

* `tapply(vector, index, function)`

```{r}
n <- 17 
fac <- factor(rep_len(1:3, n), levels = 1:5)
table(fac)
tapply(1:n, fac, sum)
```

`mapply(fun, ..., MoreArgs)`
* `fun` - function to be performed
* `...` - arguments to vectorize over 
* `MoreArgs` - list of other arguments to function, constant oens

```{r}
mapply(rep, times = 1:4, MoreArgs = list(x = 42))
```

# More tidyverse

```{r}
library(tidyverse)
data(starwars)
starwars
```

## Revision

### summarise

```{r}
starwars %>% 
  summarise(median_height = median(height),
            median_mass = median(mass))
```

### group_by

```{r}
starwars %>%
  group_by(gender) %>%
  summarise(median_height = median(height),
            median_mass = median(mass))
```

### mutate

```{r}
starwars %>%
  mutate(height_in = height / 2.54)
```

### filter

```{r}
starwars %>%
  group_by(gender) %>%
  filter(!is.na(gender)) %>%
  summarise(median_height = median(height),
            median_mass = median(mass))
```

### top_n

```{r}
starwars %>%
  group_by(gender) %>%
  top_n(1, height)
```





