---
title: "Introduction to R - Part 3 Flow - notebook"
author: "Katarzyna Kedzierska"
date: "21 August 2019"
output: html_notebook
---

# Last meeting reminder

Write a function `my_fun2` that will take two arguments - a vector and a number and:
* add the number to elemnts at odd postiions of your vector
* mutliply the elements at even positions of your vector
* raise the whole vector to the power of that number

```{r}
my_fun2 <- function(vec, numb) {
  vec2 <- vec
  vec2[1:length(vec2) %% 2 == 1] <- vec2[1:length(vec2) %% 2 == 1] + numb
  vec2[1:length(vec2) %% 2 == 0] <- vec2[1:length(vec2) %% 2 == 0] * numb
  vec2 ^ numb
}
```

```{r}
my_fun2(1:5, 2)
```

# Controlling the flow
## Logic in R

Before we will learn about the logic let's define a couple useful vectors.
```{r}
vec1 <- c(TRUE, FALSE, TRUE, TRUE)
vec2 <- rep(TRUE, 4)
```

Logical operators in R:

`TRUE`, `FALSE` are respectively equal to 1 and 0;
```{r}
TRUE + FALSE
```
```{r}
sum(vec1)
```

`R` has two operators for **and**: `&`, `&&` means both values have to be true for it to be true;

`&` performs elementwise operations.
```{r}
vec1 & vec2
```
Remember about the `R` recycling!
```{r}
vec1 & c(TRUE, FALSE)
```
In turn, `&&` only performs the comparison from left to right.
```{r}
vec1 && vec2
```
or until condition is satisfied
```{r}
TRUE && TRUE && FALSE && TRUE
```

Similarily, `R` has two opperands for **or**: `|`, `||` means either value has to be true for it to be true;  

```{r}
vec1 | vec2
```
vs.
```{r}
vec1 || vec2
```
```{r}
TRUE || FALSE || TRUE || TRUE || TRUE
```

To negate the value use `!`
```{r}
!TRUE
```
```{r}
!vec1
```

Comparisons: 
* `==` - equal to
```{r}
TRUE == FALSE
```
* `!=` - not equal to
```{r}
vec1 != vec2
```

* `<` and `<=` - smaller, smaller or equal to
```{r}
2 < 4
```

```{r}
5 <= 5
```
* `>` and `>=` - greater, greater and equal to
```{r}
5 > 1
```

```{r}
5 >= 5
```

`xor()` - returns `TRUE` when only one value equal to `TRUE`, accepts only 2 arguments;
```{r}
xor(TRUE, FALSE)
```

```{r}
xor(TRUE, TRUE)
```

`any()` - returns `TRUE` when **at least one** value equal to `TRUE`;
```{r}
any(vec1)
```

`all()` - returns `TRUE` when **all** values equal to `TRUE`;
```{r}
all(vec1)
```

`isTRUE()`, `isFALSE()` - checks if values `TRUE` or `FALSE` respectively.
```{r}
isTRUE(TRUE)
```
but, it only will report true if presented with `TRUE` value.

```{r}
isTRUE(vec2)
```

## Controlling the flow

`if()` - given certain condition, perform an action

```{r}
if (TRUE) {
  print("Action performed when TRUE.")
}
```

`else` - when statement in `if` not true

```{r}
if (FALSE) {
  print("This will not be printed.")
} else {
  print("But this will.")
}
```

Let's have some fun

```{r}
input <- readline("Hiya! What's your name? ")

if (input != "") {
  print(paste0("Hi ", input, ". Nice to meet you! :)"))
} else {
  print("Don't be shy!")
}
```

But what if we want to test multiple conditions? `else if` is the answer.

```{r}
input <- as.numeric(readline("Please, provide a number greater than 0: "))

if (input > 0) {
  print("Thanks!")
} else if (input == 0) {
  print("Zero is not greater than zero, it's equal!")
} else {
  print("C'mon, you can do better!")
}
```

**Excercise**
Write the chunk of code that will test if an number is odd (print "It's an odd number!"), divisible by 6 (print "It's a number divisible by 6!") or odd not divisible by 3 ("It's an even number, not divisible by 3."). Test the code on a random number.

```{r}
number <- sample(1:100, 1)
print(number)
# your code goes here
```

`if` works with single `TRUE` or `FALSE` value. Ther is a vectorised version though! `ifelse`!

```{r}
print(vec1)
```

```{r}
ifelse(vec1, 
       "this is what is assigned when TRUE", 
       "and this when FALSE")
```

**Excercise**
Write the chunk of code that will test if an number is odd (print "It's an odd number!"), divisible by 6 (print "It's a number divisible by 6!") or odd not divisible by 3 ("It's an even number, not divisible by 3."). Test the code on a random number. <- This, but using `ifelse`

```{r}
numbers <- sample(1:100, 5)
print(numbers)

# your code goes here
```

### Loops: for and while

Those statements allow us to repeat a certain procedure over an iteration of sorts. For example, let's say that for every odd number lower than 10 we would want to learn the value of this number taken to the power of 3. We can do it like this:

```{r}
c(1:10)[c(TRUE, FALSE)]^3
```

But, with `for` loop, we can do it like this:

```{r}
for (n in 1:10) { 
  if (n %% 2 == 1) {
    print(n ^ 3) # what will happen without print?
  }
}
```

Let's get the sum of all of the above!

```{r}
sum_of_cubes <- 0
for (n in seq(1, 10, 2)) { 
  print(n)
  print(sum_of_cubes)
  sum_of_cubes = sum_of_cubes + n ^ 3
}

print(sum_of_cubes)
```

Any alternative for the above loop?

```{r}
sum_of_cubes <- 0
for (n in seq(1, 10, 2)) { 
  sum_of_cubes = sum_of_cubes + n ^ 3
}

print(sum_of_cubes)
```

Any alternative for the above loop?

```{r}
sum_of_cubes <- sum(seq(1, 10, 2)^3)
```

`while` loop executes until it is interupted.

```{r}
i <- 10
while (i < 100) {
  print(i)
  i = i + i 
}
```

`break` will break the loop

```{r}
while (TRUE) {
  number <- sample(1:3, 1)
  if (number == 2) {
    print("Found my number!")
    break
  } else {
    print("Still not my number :(")
  }
}
```

```{r, eval = FALSE}
while (TRUE) {
  input <- as.numeric(readline("Please, provide a number greater than 0: "))
  if (input > 0) {
    print("Thanks!")
    break
  } else if (input == 0) {
    print("Zero is not greater than zero, it's equal!")
  } else {
    print("C'mon, you can do better!")
  }
}
```

`next` is a statement that allows skipping the rest of the loop.

```{r}
for (n in 1:10) {
  if (n %% 2 == 0) {
    print(paste0("Found an even number - ", n))
    next()
  }
  print(paste0("Found an odd number - ", n))
}
```


**Excercise**  
Write a loop that will print the following pattern:  
1  
22  
333  
4444    
55555  
666666  
7777777  
88888888  
999999999  
  
*Hint*: check out those functions: `rep` and `paste`.

```{r}
# your code here
```

**Excercise**  
Write a function called `fizz_buzz` that takes a number and:
* If the number is divisible by 3, it should print "Fizz".
* If it is divisible by 5, it should print "Buzz".
* If it is divisible by both 3 and 5, it should print "FizzBuzz".
* Otherwise, it should print the same number.

```{r}
fizz_buzz <- function(x) {
  if (x %% 5 == 0 & x %% 3 == 0) {
   print("FizzBuzz")
  } else if (x %% 5 == 0) {
   print("Buzz")
  } else if (x %% 3 == 0) {
   print("Fizz")
 } else {
    print(x)
  }
}

# check
fizz_buzz(9) # Fizz
fizz_buzz(25) # Buzz
fizz_buzz(15) # FizzBuzz
fizz_buzz(13) # 13
```

**Excercise**  
Write a function for checking the speed of drivers - `check_speed`. This function should have one parameter: speed.
* If speed is less than 70, it should print "Ok".  
* Otherwise, for every 5km above the speed limit (70), it should give the driver one demerit point and print the total number of demerit points. For example, if the speed is 80, it should print: "Points: 2".
* If the driver gets more than 12 points, the function should print: "License suspended"
```{r}
check_speed <- function(speed) {
  # your code goes here
}

# check
check_speed(64) # Ok
check_speed(71) # Points: 1
check_speed(83) # Points: 3
check_speed(131) # Licence suspended
```