---
title: "Introduction to R - Part 4 Data Visualisation"
author: "Katarzyna Kedzierska"
date: "4 September 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
---

```{r, include = FALSE}
knitr::opts_chunk$set(message = FALSE,
                      dpi = 300, 
                      out.height = "300px", 
                      out.width = "300px")
```

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
mono_light(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
  header_h1_font_size = "40px",
  header_h2_font_size = "30px",
  header_h3_font_size = "20px", 
  text_font_size = "20px",
  code_font_size = "15px"
)
```

# Catch up

So far we learned:

* about data types and structures;
* how to write functions;
* about the logic and the flow.

---

# Quiz

```{r}
vec1 <- c("wt", "wt", "mut1", "mut2", "mut1", "mut1", "mut2", "wt")
vec2 <- as.factor(vec1)
```

## What's the order of classes in the vec2?

---

# Quiz

```{r}
vec1 <- c("wt", "wt", "mut1", "mut2", "mut1", "mut1", "mut2", "wt")
vec2 <- as.factor(vec1)
```

## What's the order of classes in the vec2?

```{r}
vec2
```

---

# Quiz

```{r}
vec1 <- c("wt", "wt", "mut1", "mut2", "mut1", "mut1", "mut2", "wt")
vec2 <- as.factor(vec1)
```

## Is `vec1` equal to `vec2`? How to check this?

---

# Quiz

```{r}
vec1 <- c("wt", "wt", "mut1", "mut2", "mut1", "mut1", "mut2", "wt")
vec2 <- as.factor(vec1)
```

## Is `vec1` equal to `vec2`? How to check this?

```{r}
identical(vec1, vec2)
```
or

```{r}
all(vec1 == vec2) && typeof(vec1) == typeof(vec2)
```

---

# Quiz

## What will the following code return?
```{r, eval=FALSE}
x <- 10
f1 <- function(y) {
  y = x
  x + y
}
f1(8)
```

---

# Quiz

## What will the following code return?
```{r}
x <- 10
f1 <- function(y) {
  y = x
  x + y
}
f1(8)
```

---

# Quiz

## What will the following code return?
```{r, eval=FALSE}
x <- 10
f1 <- function(x) {
  function() {
    x + 10
  }
}
f1(1)()
```
---

# Quiz

## What will the following code return?
```{r}
x <- 10
f1 <- function(x) {
  function() {
    x + 10
  }
}
f1(1)()
```
---


# Why visualise data?
Let's look at this data set with 20 observations (rows) for 2 variables (columns).

```{r, echo=FALSE, include=FALSE}
library(tidyverse)
set.seed(0)
t <- sort(runif(20, 0, 2*pi))
x <- 2*cos(t)
y <- 2*sin(t)
df <- data.frame(x, y)
```
.pull-left[
```{r}
df
```
]
.pull-right[
```{r}
summary(df)
cor(df$x, df$y)
```
]

---

# Why visualise data?

Can we say more about the data now?

```{r}
ggplot(df, aes(x, y)) +
  geom_point()
```

---
However, we need to remember that there are better and worse ways to visualise data.

.pull-left[
```{r}
ggplot(df, aes(x, y)) +
  geom_point()
```
]
.pull-right[
```{r}
df %>% 
  mutate(index = 1:n()) %>%
  gather(variable, value, -index) %>%
  ggplot(., aes(index, value, color = variable)) +
  geom_line()
```
]
.footnote[The example data set and the bad/good plot idea came from [Kaspar Martens](https://kasparmartens.rbind.io/) [slides](https://github.com/kasparmartens/2018_12_bham/blob/master/2018_10_18_slides.pdf) which were inspired by the talk "How Humans See Data" by John Rauser]

---

# ggplot2

Please refer to the cheat sheets. This is a great resource to begin your adventure with `ggplot2`. We will discover the `ggplot2` package, and some of the `tidyverse` functions as we will go on. 

.pull-left[
```{r, echo=FALSE}
plot(PlantGrowth$group, 
     PlantGrowth$weight, 
     main = "Ugly, but easy, plot")
```
]
.pull-right[
```{r, echo=FALSE}
library(ggsci)
PlantGrowth %>%
  ggplot(., aes(group, weight)) +
  geom_jitter(aes(color = group)) +
  geom_violin(alpha = 0.2, aes(fill = group)) +
  geom_boxplot(width = 0.1) +
  theme_classic() +
  scale_color_futurama() +
  scale_fill_futurama() +
  labs(x = "Group", color = "Group", 
       fill = "Group", y = "Weight", title = "Pretty plot") +
  theme(legend.position = "bottom")
```
]

---

## Simple plot
Let's start with the basics though. We will create our first plot with the following data.

```{r}
df <- data.frame(x = -100:100,
                 y = (-100:100)^2)
summary(df)
```

---
## Simple plot

```{r}
plt <- ggplot()
plt
```
---

## Simple plot

```{r}
plt <- ggplot(df)
plt
```
---

## Simple plot

```{r}
plt <- ggplot(df, aes(x, y))
plt
```

---

## Simple plot

```{r}
plt <- ggplot(df, aes(x, y))
plt + geom_point()
```

---
## Excercise 
Given `iris` data set:

```{r}
head(iris)
```

```{r}
summary(iris)
```

---
Plot the following figure. (We want to show the relationship between Sepal Length and Sepal Width with respect to the Specie.)
```{r, echo=FALSE}
iris %>%
  ggplot(., aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point() +
  geom_smooth()
```

**Hint:** Use your cheatsheets!

---

# Which one is better?
.pull-left[
```{r, echo = FALSE}
iris %>%
  ggplot(., aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point() +
  geom_smooth()
```
]
.pull-right[
```{r, echo = FALSE}
iris %>%
  ggplot(., aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~Species)
```
]

---

## Faceting

`facet_wrap()` allows to plot subplots

```{r}
iris %>%
  ggplot(., aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~Species, )
```

