---
title: "Introduction to R - Part 7"
author: "Katarzyna Kedzierska"
date: "2 October 2019"
output:
  html_document:
    df_print: paged
---

```{r setup}
needed_packages <- c("tidyverse", "ggsci")
for (pkg in needed_packages) {
  if (!require(pkg, quietly = TRUE, character.only = TRUE)) {
    print(paste0("WARNING! Installing ", pkg))
    install.packages(pkg)
  }
  library(pkg, character.only = TRUE)
}
```

# More tidyverse

We will be using the `starwars` data. Spent a momemnt getting to know this data.

```{r}
data(starwars)
starwars %>% 
  head()
```

Look at its description
```{r}
?starwars
```

and its summary.
```{r}
summary(starwars[,1:10])
```

# Revision

## summarise

```{r}
# this will not work because of missing data
# starwars %>% 
#   summarise(median_height = quantile(height, 0.9),
#             median_mass = quantile(mass, 0.9))
```

Be aware of how you deal with missing data. You may be tempted, and sometise it's understandable choice, to use simply use `na.rm = TRUE`. Be warned though, given calculation can be refferred to only subset of your data (i.e. the one that is not missing).

Be also causious, as the following:

```{r}
starwars %>% 
  summarise(median_height = quantile(height, 0.9, na.rm = TRUE),
            median_mass = quantile(mass, 0.9, na.rm = TRUE))
```

```{r}
starwars %>% 
  filter(!is.na(height), !is.na(mass)) %>%
  summarise(median_height = quantile(height, 0.9),
            median_mass = quantile(mass, 0.9))
```
```{r}
starwars %>% 
  na.omit() %>%
  summarise(median_height = quantile(height, 0.9),
            median_mass = quantile(mass, 0.9))
```
will give 3 different results. Why?

* `na.rm = TRUE` will result in removal of only the missing value for a cerain calculation;   
* `filter(!is.na(height), !is.na(mass))` will remove all rows with `NA` in the height column or mass;
* `na.omit()` will remove rows with at least one `NA`.

## group_by

```{r}
starwars %>%
  group_by(gender) %>%
  summarise(median_height = median(height, na.rm = TRUE),
            median_mass = median(mass, na.rm = TRUE))
```

## mutate

```{r}
starwars %>%
  mutate(height_in = height / 2.54) %>%
  select(name, height, height_in) %>%
  head()
```

## filter

```{r}
starwars %>%
  group_by(gender) %>%
  filter(!is.na(gender), !is.na(height), !is.na(mass)) %>%
  summarise(median_height = median(height),
            median_mass = median(mass))
```

## top_n

```{r}
starwars %>%
  group_by(gender) %>%
  top_n(1, height)
```

## gather

```{r}
starwars %>%
  select(name, height, mass) %>%
  arrange(name) %>%
  head()
```


```{r}
starwars %>%
  select(name, height, mass) %>%
  gather(key = variable, 
         value = value, 
         height, mass) %>%
  arrange(name) %>%
  head()
```


## spread

```{r}
starwars %>%
  select(name, height, mass) %>%
  gather(key = variable, 
         value = value, 
         height, mass) %>%
  arrange(name) %>%
  head()
```


```{r}
starwars %>%
  select(name, height, mass) %>%
  gather(key = variable, 
         value = value, 
         height, mass) %>%
  spread(variable, value) %>%
  arrange(name) %>%
  head()
```


## join

```{r, echo=FALSE}
tmp_df <- data.frame(gender = c("female", "hermaphrodite", "none", NA),
                     genedr_modified = c("Female", "Hermaphrodite", 
                                         "None", "No information"))
```

```{r}
tmp_df
```

```{r, warning=FALSE, message=FALSE}
starwars %>%
  select(name, gender) %>%
  right_join(tmp_df) %>%
  arrange(name) %>%
  head()
```

```{r, warning=FALSE, message}
starwars %>%
  select(name, gender) %>%
  full_join(tmp_df) %>%
  arrange(name) %>%
  head()
```

# Excercises

1. Which specie is the heaviest?  
2. Who's the tallest?  
3. Who is the thinnest? (lowest mass to height ratio)  
4. Who is an outlier on the height to mass ratio plot below. 
5. Who has appeared in all movies? Who has appeared in the least movies?  
6. List the most represented species from each homeworld.  

```{r}
starwars %>% 
  mutate(out = mass>1000) %>%
  replace_na(list(out = FALSE)) %>%
  ggplot(., aes(height, mass, color = out, shape = out)) +
  geom_point(size = 5) +
  scale_color_aaas() +
  theme(legend.position = "none")
```

1 - Which specie is the heaviest?  

```{r}
starwars %>%
  group_by(species) %>%
  summarise(med_mass = median(mass, na.rm = TRUE)) %>%
  top_n(1, med_mass)
```

2 - Who's the tallest?  

```{r}
starwars %>% 
  top_n(1, height)
```

3 - Who is the thinnest? (lowest mass to height ratio)  

```{r}
starwars %>%
  mutate(ratio = mass / height) %>%
  top_n(-1, ratio)
```

4 - Who is an outlier on the height to mass ratio plot below. 


```{r}
starwars %>%
  select(name, species, mass) %>%
  top_n(1, mass)
```

5 - Who appeared in all movies?  

```{r}
starwars %>%
  unnest(films) %>%
  group_by(name) %>%
  summarise(n_films = n()) %>%
  ungroup() %>%
  top_n(1, n_films)
```

Who appeared in the least number of movies? 

```{r}
starwars %>%
  unnest(films) %>%
  group_by(name) %>%
  summarise(n_films = n()) %>%
  ungroup() %>%
  top_n(-1, n_films)
```

6 - List the most represented species from each homeworld.  

```{r}

starwars %>%
  group_by(homeworld, species) %>%
  summarise(n_beings = n()) %>%
  group_by(species) %>%
  top_n(1, n_beings)

```

## More excercises 

Column `starships` stores information of starships given charater piloted. It stores the names of those starships as a `list`. Firstly, create a new column `n_pilots` which stores number of pilots of each straship. Secondly, list all the characters that piloted the same starship together with that starship name.

```{r}
starwars %>%
  unnest(starships) %>%
  group_by(starships) %>%
  mutate(n_pilots = n())  %>%
  filter(n_pilots > 1) %>%
  select(name, starships, n_pilots) %>%
  arrange(-n_pilots, starships, name) %>%
  ##ungroup() %>%
  nest(name)


starwars %>%
  unnest(starships) %>%
  group_by(starships) %>%
  mutate(n_pilots = n()) %>%
  filter(n_pilots > 1) %>%
  select(name, starships, n_pilots) %>%
  arrange(-n_pilots, starships, name) %>%
  nest(name)
```
---

Do the same for the `vehicles` column.

```{r}
starwars %>%
  unnest(vehicles) %>%
  group_by(vehicles) %>%
  mutate(n_pilots = n())  %>%
  filter(n_pilots > 1) %>%
  select(name, vehicles, n_pilots) %>%
  arrange(-n_pilots, vehicles, name)
```
